---
title: "bike_share_company"
author: "Anord Nkuba"
date: "2023-04-15"
output: html_document
---

#load packages
```{r}
library(tidyverse)
library(lubridate)
library(janitor)
library(data.table)
library(readr)
library(psych)
library(hrbrthemes)
library(ggplot2)
```


Total no. of customers
```{r}
print(table(clean_trip_final9$costumer_type))
```

#total number of casual and member customer from January to December 2022
```{r}
total_customer <- (table(clean_trip_final9$costumer_type))
```

Data visualization total number of casual and member customer from January to December 2022
```{r}
ggplot(data = data.frame(total_customer), aes(x = Var1, y = Freq)) + geom_bar(stat = "identity") + 
  labs(title="total number of casual&member riders from Jan-Dec", x="customer type", y="Orders Frequency")
```


#Total rides for each customer type in minutes
```{r}
total_ride <- setNames(aggregate(ride_length ~ costumer_type, clean_trip_final9, sum), c("customer_type", "total_ride_min"))
print(total_ride)
```

#data visualization of total rides by each customer type
```{r}


ggplot(total_ride, aes(x=customer_type, y=total_ride_min)) + 
  geom_bar(stat="identity", fill="grey") + 
  labs(title="Total Ride Length by Customer Type", x="Customer Type", y="Total Ride Length (mins)")
```

#Differences between members and casual riders in terms of length of ride
```{r}
print(clean_trip_final9 %>% 
       group_by(costumer_type) %>% 
       summarise(min_length_mins = min(ride_length), max_length_min = max(ride_length),
                 median_length_mins = median(ride_length), mean_length_min = mean(ride_length)))
```

#data visualization average ride length between casual and member riders from Jan to Dec
```{r}
mean_ride_length <- clean_trip_final9 %>% group_by(costumer_type) %>% summarise(mean_length_min = mean(ride_length))

ggplot(data = mean_ride_length, aes(x = costumer_type, y = mean_length_min)) + geom_bar(stat = "identity")

```


#Average ride_length for users by day_of_week and Number of total rides by day_of_week
```{r}
print(clean_trip_final9 %>% 
       group_by(week_day) %>% 
       summarise(Avg_length = mean(ride_length),
                 number_of_ride = n()))
```

#Average ride_length by month
```{r}
print(clean_trip_final9 %>% 
       group_by(month) %>% 
       summarise(Avg_length = mean(ride_length),
                 number_of_ride = n()))
```

#Average ride length comparison by each week day according to each customer type
```{r}
print(aggregate(clean_trip_final9$ride_length ~ clean_trip_final9$costumer_type + 
                 clean_trip_final9$week_day, FUN = mean))
```

#Average ride length comparison by each month according to each customer type
```{r}
print(aggregate(clean_trip_final9$ride_length ~ clean_trip_final9$costumer_type + 
                 clean_trip_final9$month, FUN = mean))
```

#Analyze rider length data by customer type and weekday
```{r}
print(clean_trip_final9 %>% 
       group_by(costumer_type, week_day) %>% 
       summarise(number_of_ride = n(),
                 avgerage_duration = mean(ride_length),
                 median_duration = median(ride_length),
                 max_duration = max(ride_length),
                 min_duration = min(ride_length)))

```

#Analyze rider length data by customer type and month
```{r}
print(clean_trip_final9 %>% 
       group_by(costumer_type, month) %>% 
       summarise(nummber_of_ride = n(),
                 average_duration = mean(ride_length),
                 median_duration = median(ride_length),
                 max_duration = max(ride_length),
                 min_duration = min(ride_length)))
```






